## Lecture 9: Calibrating experts {#calibration_lecture} 
<!-- reference with [Calibrating experts](#calibration_lecture) -->

Welcome to lecture 9 of **Decision Analysis and Forecasting for Agricultural Development**. Feel free to bring up any questions or concerns in the Slack or to [Dr. Cory Whitney](mailto:cory.whitney@uni-bonn.de?subject=[Lecture_4]%Calibrating%Experts%20Lecture), [Dr. Prajna Kasargodu Anebagilu](mailto:pkasargo@uni-bonn.de?subject=[Lecture_7]%20Decision%20Analysis%20Lecture) or the course tutor.

The following videos tell you something about a method which is frequently used in our group to get reliable values for the input tables from our stakeholders. Please watch the videos and answer the questions that follow. 
In this class, you will learn about different techniques you can use to improving your estimations. And during Seminar 7, you will get to test this out using our Calibration App. 

### What is calibration training?

<iframe width="560" height="315" src="https://www.youtube.com/embed/DYAv-LBZaIU" frameborder="0" allowfullscreen></iframe>

```{r question-explaining-calibration-training-part-01-01, echo=FALSE}
question("What is the intention of decision analysts in measuring things?",
  answer("They want to get paid for it."),
  answer("They want to get precise information."),
  answer("They follow what the research project leader tells them to do."),
 answer("They want to reduce their uncertainty.", correct = TRUE),
 answer("They are inspired by Douglas Hubbard to try and measure everything."), 
  incorrect = "Watch [the video](https://youtu.be/DYAv-LBZaIU) especially the part
 4:30-5:30 and try again.",
allow_retry = TRUE
)
```


```{r question-explaining-calibration-training-part-01-02, echo=FALSE}
question("What is calibration?",
answer("Calibration is a measuring process.", correct = TRUE),
answer("Calibration is a test of measuring devices to make sure they are measuring accurately."),
answer("Calibration tells us which measuring device not to use."),
answer("Calibration tells us the deviation of a measuring device.", correct = TRUE),
answer("Calibration helps our experts to get a feeling about their uncertainty.", correct = TRUE),
  incorrect = "Watch [the video](https://youtu.be/pPIs7MwgMe0) especially the part
12:35-14:15 and try again.",
allow_retry = TRUE
)
```

### Why do we apply calibration training?

<iframe width="560" height="315" src="https://www.youtube.com/embed/JqOX3cHa8AI" title="Calibration Training Part 2" frameborder="0" allowfullscreen></iframe>

```{r question-explaining-calibration-training-part-02-01, echo=FALSE}
question("Why do we apply the calibration training process with experts?",
answer("To tell them more about the desired field of expertise."),
answer("To train them about their own biases.", correct = TRUE),
answer("To get to know them better."),
answer("To teach them procedures how to improve the estimation performance.", correct = TRUE),
answer("To let them interact with a wheel of fortune."),
answer("To slow down their answering behaviour and reduce the impact of the affective system.", correct = TRUE),
  incorrect = "Watch [the video](https://youtu.be/pPIs7MwgMe0) and try again.",
allow_retry = TRUE
)
```

```{r question-explaining-calibration-training-part-02-02, echo=FALSE}
question("What is the meaning of 'Anchoring'?",
answer("Achoring is the influence of multicolored wheels on estimates."),
answer("Anchoring is specifically about ships and is unrelated to the calibration topic."),
answer("Anchoring is a kind of bias that may influence you to give estimates around a value you have seen or heard, but is not necessarily related to the variable you are estimating.", correct = TRUE),
  incorrect = "Watch [the video](https://youtu.be/pPIs7MwgMe0) and try again.",
allow_retry = TRUE
)
```



### Group discussion reading: 

- **Chapter 1.** Kahneman, Daniel, and Patrick Egan. Thinking, Fast and Slow. Vol. 1. New York: Farrar, Straus and Giroux, 2011.

### Bonus Material

Watch the video and do the awareness test!

<iframe width="560" height="315" src="https://www.youtube.com/embed/UfA3ivLK_tE" frameborder="0" allowfullscreen></iframe>

Have a look at Terry Heick's [Cognitive Bias codex](https://www.teachthought.com/critical-thinking/the-cognitive-bias-codex-a-visual-of-180-cognitive-biases/). 