### Plot the impact pathway with mermaid

In this example, we show how we can transform a simple impact pathway into code for estimating profit distributions.

We use the `mermaid` function from the `DiagrammeR` library to create the graphical impact pathway [@R-DiagrammeR].

Run the code below and then add an additional cost to the graph called `Management cost` (try to do this on your own but feel free to look to the `solution` to see one way to do this).

```{r graph-mermaid, exercise=TRUE}
mermaid("graph LR
        Y(Yield)-->I(Income); linkStyle 0 stroke:green, stroke-width:1.5px
        M(Market price)-->I; linkStyle 1 stroke: green, stroke-width:1.5px
        I-->F(Final result); linkStyle 2 stroke: green, stroke-width:1.5px
        CL(Labor cost)-->F; linkStyle 3 stroke: red, stroke-width:1.5px")
```

```{r graph-mermaid-solution}
mermaid("graph LR
        Y(Yield)-->I(Income); linkStyle 0 stroke:green, stroke-width:1.5px
        M(Market price)-->I; linkStyle 1 stroke: green, stroke-width:1.5px
        I-->F(Final result); linkStyle 2 stroke: green, stroke-width:1.5px
        CL(Labor cost)-->F; linkStyle 3 stroke: red, stroke-width:1.5px
        CM(Management cost)-->F; linkStyle 4 stroke: red, stroke-width:1.5px")
```

The `LR` argument in the `mermaid` function specifies that the plot will go from left to right. Try and change this to `TD`.

Use the `style` options in `mermaid` to change the arrow widths to `2px` and the node colors to red for costs and green for benefits. You will need to break the line and put the `linkStyle` on a new line to add the color to the node.

```{r graph-mermaid-direction, exercise=TRUE}
mermaid("graph LR
        Y(Yield)-->I(Income); linkStyle 0 stroke:green, stroke-width:1px
        M(Market price)-->I; linkStyle 1 stroke: green, stroke-width:1px
        I-->F(Final result); linkStyle 2 stroke: green, stroke-width:1px
        CL(Labor cost)-->F; linkStyle 3 stroke: red, stroke-width:1px
        CM(Management cost)-->F; linkStyle 4 stroke: red, stroke-width:1px")
```

```{r graph-mermaid-direction-solution}
mermaid("graph TB
        Y(Yield)-->I(Income); style I fill:green
        linkStyle 0 stroke:green, stroke-width:2px
        M(Market price)-->I; linkStyle 1 stroke: green, stroke-width:2px
        I-->F(Final result); linkStyle 2 stroke: green, stroke-width:2px
        CL(Labor cost)-->F; style CL fill:red
        linkStyle 3 stroke: red, stroke-width:2px
        CM(Management cost)-->F; style CM fill:red
        linkStyle 4 stroke: red, stroke-width:2px")
```

That was just one of many ways to generate impact pathways. To see more options see the [Decision Analysis Overview](#decision_analysis) lecture materials.
