## Seminar 10: Model functions {#model_functions} 
<!-- [Model functions](#model_functions) -->

Welcome to tenth seminar of **Decision Analysis and Forecasting for Agricultural Development**.Feel free to bring up any questions or concerns in the Slack or to [Dr. Cory Whitney](mailto:cory.whitney@uni-bonn.de?subject=[Seminar_10]%20Decision%20Analysis%20Lecture) or the course tutor.

In this seminar we will go over some of the functions that are useful for building parts of our models in the `decisionSupport` package. 

### The value varier function

Many variables vary over time and it may not be desirable to ignore this variation in time series analyses. The `vv` function produces time series that contain variation from a specified mean and a desired coefficient of variation. A trend can be added to this time series.

The arguments for the `vv` function include: 

- `var_mean`, which is the mean of the variable to be varied
- `var_CV`, which is the desired coefficient of variation (in percent)
- `n`, which is the integer; number of values to produce
- `distribution`, which is the probability distribution for the introducing variation. This is currently only implemented for normal distributions
- `absolute_trend`, which is the absolute increment in the `var_mean` in each time step. Defaults to `NA`, which means no such absolute value trend is present. If both absolute and relative trends are specified, only original means are used
- `relative_trend`, which is the relative trend in the `var_mean` in each time step (in percent). Defaults to `NA`, which means no such relative value trend is present. If both absolute and relative trends are specified, only original means are used
- `lower_limit`, which is the lowest possible value for elements of the resulting vector
- `upper_limit`, which is the upper possible value for elements of the resulting vector

Note that only one type of trend can be specified. If neither of the trend parameters are NA, the function uses only the original means

The function produces a vector of `n` numeric values, representing a variable time series, which initially has the mean `var_mean`, and then increases according to the specified trends. Variation is determined by the given coefficient of variation `var_CV`.

Create a vector with the `vv` function and plot with the base R `plot` function. 

```{r reports-seminar-vv, exercise=TRUE}
# reduce var_CV to 5 and change n to 40. Name the output 'valvar' and plot with base R  
vv(var_mean = 100, 
   var_CV = 10, 
   n = 30)

```

```{r reports-seminar-vv-solution}
valvar <- vv(var_mean = 100, 
             var_CV = 5, 
             n = 40)

plot(valvar)
```

Use the `absolute_trend` argument and plot with the base R `plot` function. 

```{r reports-seminar-absolute_trend, exercise=TRUE}
# reduce var_mean to 50 and make absolute_trend 10. Name the output 'valvar' and plot with base R 
vv(var_mean = 100, 
           var_CV = 10, 
           n = 30, 
           absolute_trend = 5)

```

```{r reports-seminar-absolute_trend-solution}
valvar <- vv(var_mean = 50, 
           var_CV = 10, 
           n = 30, 
           absolute_trend = 10)

plot(valvar)
```

Use the `relative_trend` argument and plot with the base R `plot` function. 

```{r reports-seminar-relative_trend, exercise=TRUE}
# reduce var_CV to 5 and change n to 40. Name the output 'valvar' and plot with base R 
vv(var_mean = 100, 
             var_CV = 10, 
             n = 30, 
             relative_trend = 5)

```

```{r reports-seminar-relative_trend-solution}
valvar <- vv(var_mean = 100, 
             var_CV = 5, 
             n = 40, 
             relative_trend = 5)

plot(valvar)
```

### Simulate occurrence of random events

In many simulations, certain events can either occur or not, and values for dependent variables can depend on which of the cases occurs. This function randomly simulates whether events occur and returns output values accordingly. The outputs can be single values or series of values, with the option of introducing artificial variation into this dataset.

The arguments for the `chance_event` function include: 

- `chance`, which is the probability that the risky event will occur (between 0 and 1)
- `value_if`, which is the output value in case the event occurs. This can be either a single numeric value or a numeric vector. Defaults to 1.
- `value_if_not`, which is the output value in case the event does not occur. This can be either a single numeric value or a numeric vector. If it is a vector, it must have the same length as `value_if`
- `n`, which is the number of times the risky event is simulated. This is ignored if `length(value_if)>1`.
- `CV_if`, which is the coefficient of variation for introducing randomness into the value_if data set. This defaults to 0 for no artificial variation. See documentation for the `vv` function for details.
`CV_if_not`, which is the coefficient of variation for introducing randomness into the value_if_not data set. This defaults to the value for `CV_if.` See documentation for the `vv` function for details.
- `one_draw`, which is the boolean coefficient indicating if event occurrence is determined only once (`TRUE`) with results applying to all elements of the results vector, or if event occurrence is determined independently for each element (`FALSE` is the default). 

The `chance_event` function provides a numeric vector of length `n`, containing outputs of a probabilistic simulation that assigns `value_if` if the event occurs, or `value_if_not` if is does not occur (both optionally with artificial variation).

```{r reports-seminar-chance-event, exercise=TRUE}
# decrease the chance and value_if by half and repeat 20 times. Name the output 'chancevar' and plot with base R 
chance_event(chance = 0.5, 
             value_if = 6, 
             n = 10)
```

```{r reports-seminar-chance-event-solution}
chancevar <- chance_event(chance = 0.25, 
             value_if = 3, 
             n = 20)

plot(chancevar)
```

Use the `value_if_not` and `CV_if` arguments. 

```{r reports-seminar-chance-event-2, exercise=TRUE}
# make the chance 10 percent, the value_if 5 and the value_if_not 20, repeat 100 times and reduce the coefficient of variation by half. Name the output 'chancevar' and plot with base R.
chance_event(chance = 0.5,
             value_if = 1,
             value_if_not = 5,
             n = 10,
             CV_if = 20)
```

```{r reports-seminar-chance-event-2-solution}
chancevar <- chance_event(chance = 0.1,
             value_if = 5,
             value_if_not = 20,
             n = 100,
             CV_if = 10)

plot(chancevar)
```

### Gompertz function yield prediction for perennials

Yields of trees or other perennial plants have to be simulated in order to predict the outcomes of many interventions. Unlike annual crops, however, trees normally yield nothing for a few years after planting, following which yields gradually increase until they reach a tree-specific maximum. This is simulated with this function, which assumes that a Gompertz function is a good way to describe this (based on the general shape of the curve, not on extensive research...). The function assumes that yields remain at the maximum level, once this is reached. For long simulations, this may not be a valid assumption! The function parameters are estimated based on yield estimates for two points in time, which the user can specify. They are described by a year number and by a percentage of the maximum yield that is attained at that time.

The arguments for the `gompertz_yield` function include:

- `max_harvest`, which is the maximum harvest from the tree (in number of fruits, kg or other units)
- `time_to_first_yield_estimate`, which is the year (or other time unit) number, for which the first yield estimate is provided by `first_yield_estimate_percent`
- `time_to_second_yield_estimate`, which is the year (or other time unit) number, for which the second yield estimate is provided by `second_yield_estimate_percent`
- `first_yield_estimate_percent` percentage of the maximum yield that is attained in the year (or other time unit) given by `time_to_first_yield_estimate`
- `second_yield_estimate_percent` percentage of the maximum yield that is attained in the year (or other time unit) given by `time_to_second_yield_estimate`
- `n_years`, which is the number of years to run the simulation
- `var_CV`, which is the coefficient indicating how much variation should be introduced into the time series. If this is one numeric value, then this value is used for all variables. The default is 0, for a time series with no artificially introduced variation. See description of the `vv` function for more details on this.
- `no_yield_before_first_estimate`, which is the boolean variable indicating whether yields before the time unit indicated by `time_to_first_yield_estimate`should be 0.

The function provides a vector of `n_years` numeric values, describing the simulated yield of the perennial. This starts at 0 and, if the simulation runs for a sufficient number of years, approaches `max_harvest.` If `var_CV>0`, this time series includes artificial variation.

```{r reports-seminar-gompertz_yield, exercise=TRUE}
# create a vector where the maximum harvest is 500, which is achieved in 10 years (i.e. 100% by the second yield estimate)
gompertz_yield(max_harvest = 1000,
               time_to_first_yield_estimate = 5,
               first_yield_estimate_percent = 10,
               time_to_second_yield_estimate = 15,
               second_yield_estimate_percent = 90,
               n_years = 30)
```

```{r reports-seminar-gompertz_yield-solution}
gompertz_yield(max_harvest = 500,
               time_to_first_yield_estimate = 5,
               first_yield_estimate_percent = 10,
               time_to_second_yield_estimate = 10,
               second_yield_estimate_percent = 100,
               n_years = 30)
```

In the seminar you will provide the updated version of your decision model. 

### Bonus: Adding correlation to the model {#correlation}

To add correlation to your model visit the [Adding correlations to Decision Analysis models](http://htmlpreview.github.io/?https://github.com/CWWhitney/Decision_Analysis_Course/blob/main/correlations/correlations_example.html) vignette.