# Seminar 3: Model Programming

Welcome to the third seminar of **Decision Analysis and Forecasting for Agricultural Development**. ... . Feel free to bring up any questions or concerns in the Slack or to [Dr. Cory Whitney](mailto:cory.whitney@uni-bonn.de?subject=[Seminar_4]%20Decision%20Analysis%20Lecture) or the course tutor.

Causal model / diagram / Impact pathway / Theory of change

-	R packages decisionSupport (walk through a vignette or two)

# decisionSupport

We load the library
```{r}
library(decisionSupport)
```

Here is a simple model
```{r}
example_decision_model <- function(x, varnames)
{
  profit <- benefits-costs
  
  final_profits <- profit+500
  
  return(final_profits)
  
}

decisionSupport(inputFilePath = "data/example_decision_inputs.csv", 
                outputPath = "data/outputs", 
                write_table = TRUE, 
                welfareFunction = example_decision_model, 
                numberOfModelRuns = 10000, 
                functionSyntax = "plainNames")
```

-	RSTAN, Other modeling programs - Betancourt’s work (selected blogs, talks, git repos, Stan, HCMC)
-	Yihui (markdown etc.)
-	Assignments: 
-	1. “episode 3, with Colin Carroll. He mentions Tom Rainforth's thesis.”  – Alex Andorra (about his interview with Colin Carol on the podcast Learning Bayesian Statistics)
-	Rainforth, Tom. “Automating Inference, Learning, and Design Using Probabilistic Programming.” Doctor of Philosophy, University of Oxford, 2017.
