## Seminar 8: Models {#model_share_seminar} 
<!-- reference with [Models](#model_share_seminar) -->

Welcome to eighth seminar of **Decision Analysis and Forecasting for Agricultural Development**.Feel free to bring up any questions or concerns in the Slack or to [Dr. Cory Whitney](mailto:cory.whitney@uni-bonn.de?subject=[Seminar_4]%20Decision%20Analysis%20Lecture) or the course tutor.

### Decision analysis with the decisionSupport package

As we have discussed in previous lectures, when deciding on intervention in complex systems, many of the variables decision makers need to consider are difficult or impossible to precisely quantify. The `decisionSupport` package [@R-decisionSupport] can make use of uncertainty around these relationships and implement a Monte Carlo simulation, which generates a large number of plausible system outcomes. The model procedure draws random numbers for each input variable, which are drawn from specified probability distributions. The method requires two inputs to the `mcSimulation` function:

1)	an R function that predicts decision outcomes based on the variables named in a separate data table. This R function is customized by the user to address a particular decision problem.  

2)	an input table (normally in `comma separated values` format, ending in `.csv`) specifying the names and probability distributions for all variables used in the decision model. These distributions aim to represent the full range of possible values for each component of the model.  

The `mcSimulation` function from the `decisionSupport` package conducts a Monte Carlo analysis with repeated model runs based on probability distributions for all uncertain variables. The data table and model are customized to fit the particulars of a specific decision.

### Building a decision model as an R function

We have already built some simple models in the [Decision Models](#decision_models) lecture and the [Model programming](#model_programming) seminar. Now let's look at a slightly more comprehensive decision model of sedimentation management strategies for a reservoir in the Upper Volta River Basin of Burkina Faso [@lanzanova_improving_2019]. The reservoirs in the Upper Volta have multiple benefits for rural communities and are important for food security and livelihoods. Sedimentation is a major issue, requiring efficient sedimentation management for which there are many options each fraught with uncertainty and risk. We will walk through a simplified version of the model by @lanzanova_improving_2019. We will use various functions from the `decisionSupport` library. 

First we generate a model as a function. We use the `decisionSupport` functions `vv()` to produce time series with variation from a pre-defined mean and coefficient of variation, `chance_event()` to simulate whether events occur and `discount()` to discount values along a time series.

```{r models-seminar-burkina-model, exercise=TRUE}
example_decision_function <- function(x, varnames){
  
  # calculate ex-ante risks: impact the implementation of interventions ####
  intervention_NonPopInvolvEvent <-
    chance_event(intervention_NonPopInvolv, 1, 0, n = 1)
  
  # pre-calculate common random draws for all intervention model runs ####
  
  # profits from Tropical Livestock Units (TLU)
  TLU <- vv(TLU_no_intervention, var_CV, n_years)
  TLU_profit <- vv(profit_per_TLU, var_CV, n_years)
  
  # benefits of fruit
  precalc_intervention_fruit_benefits <-
    vv(intervention_fruit_area_ha, var_CV, n_years) *
    vv(intervention_fruit_yield_t_ha, var_CV, n_years) *
    vv(intervention_fruit_profit_USD_t, var_CV, n_years)
  
  # benefits of vegetables
  precalc_intervention_vegetable_benefits <-
    vv(intervention_vegetable_area_ha, var_CV, n_years) *
    vv(intervention_vegetable_yield_t_ha, var_CV, n_years) *
    vv(intervention_vegetable_profit_USD_t, var_CV, n_years)
  
  # benefits of rain-fed crops
  precalc_intervention_rainfed_crop_benefits <-
    vv(intervention_rainfed_crop_area_ha, var_CV, n_years) *
    vv(intervention_rainfed_crop_yield_t_ha, var_CV, n_years) *
    vv(intervention_rainfed_crop_profit_USD_t, var_CV, n_years)
  
  #  Intervention ####
  
  for (decision_intervention_strips in c(FALSE,TRUE))
      {
      
  if (decision_intervention_strips)
  {
    intervention_strips <- TRUE
    intervention_strips_PlanningCost <- TRUE
    intervention_strips_cost <- TRUE
  } else
  {
    intervention_strips <- FALSE
    intervention_strips_PlanningCost <- FALSE
    intervention_strips_cost <- FALSE
  }
  
  if (intervention_NonPopInvolvEvent) {
    intervention_strips <- FALSE
    intervention_strips_cost <- FALSE
  }
  
  # Costs ####
  if (intervention_strips_cost) {
    cost_intervention_strips <-
      intervention_adaptation_cost + 
      intervention_tech_devices_cost + 
      intervention_nursery_cost +
      intervention_wells_cost +
      intervention_training_cost + 
      intervention_mngmt_oprt_cost + 
      intervention_mngmt_follow_cost +
      intervention_mngmt_audit_cost
  } else
    cost_intervention_strips <- 0
  
  if (intervention_strips_PlanningCost) {
    plan_cost_intervention_strips <-
      intervention_communication_cost + intervention_zoning_cost
  } else
    plan_cost_intervention_strips <- 0
  
  maintenance_cost <- rep(0, n_years)
  
  if (intervention_strips)
    maintenance_cost <-
    maintenance_cost + vv(maintenance_intervention_strips, 
                          var_CV, n_years)
  
  intervention_cost <- maintenance_cost
  intervention_cost[1] <-
    intervention_cost[1] + 
    cost_intervention_strips + 
    plan_cost_intervention_strips

  
  # Benefits from  cultivation in the intervention strips ####
  
  intervention_fruit_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_fruit_benefits
  intervention_vegetable_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_vegetable_benefits
  intervention_rainfed_crop_benefits <-
    as.numeric(intervention_strips) * precalc_intervention_rainfed_crop_benefits
  
  # Total benefits from crop production (agricultural development and riparian zone) ####
  crop_production <-
    intervention_fruit_benefits +
    intervention_vegetable_benefits +
    intervention_rainfed_crop_benefits
  
  # Benefits from livestock ####
  # The following allows considering that intervention strips may
  # restrict access to the reservoir for livestock.
  
  if (intervention_strips)
    TLU_intervention <-
    TLU * (1 + change_TLU_intervention_perc / 100)
  else
    TLU_intervention <- TLU
  
  if (decision_intervention_strips){
    livestock_benefits <- TLU_intervention * TLU_profit
    total_benefits <- crop_production + livestock_benefits
    net_benefits <- total_benefits - intervention_cost
    result_interv <- net_benefits}
  
  
  if (!decision_intervention_strips){
    livestock_benefits <- TLU_no_intervention * TLU_profit
    total_benefits <- livestock_benefits
    net_benefits <- total_benefits - intervention_cost
    result_n_interv <- net_benefits}
  
    } #close intervention loop bracket

NPV_interv <-
  discount(result_interv, discount_rate, calculate_NPV = TRUE)

NPV_n_interv <-
  discount(result_n_interv, discount_rate, calculate_NPV = TRUE)

# Beware, if you do not name your outputs 
# (left-hand side of the equal sign) in the return section, 
# the variables will be called output_1, _2, etc.

return(list(Interv_NPV = NPV_interv,
            NO_Interv_NPV = NPV_n_interv,
            NPV_decision_do = NPV_interv - NPV_n_interv,
            Cashflow_decision_do = result_interv - result_n_interv))
}

```

### The input table

The second thing we need for the model is the input table describing the uncertainty in our variables. A draft of this has been generated for you already. You can get the input table from this Github repository by using the `read_csv` function from the `readr` library [@R-readr] and `url` function from base R. You can also download [the file](https://raw.githubusercontent.com/CWWhitney/Decision_Analysis_Course/main/data/example_input_table.csv) directly from GitHub if you want. You will need to use the 'save as' option for the webpage so that you can have it stored locally as a `.csv` file on your machine.

```{r models-seminar-getdata, exercise=TRUE}
library(readr)

example_input_table = "https://raw.githubusercontent.com/CWWhitney/Decision_Analysis_Course/main/data/example_input_table.csv"

input_table <- read_csv(url(example_input_table))
```

Want to have a look? Load the data using the `read.csv` function. 

```
example_data <- read.csv("example_input_table.csv")
```

Notice that the data has seven columns. Looking at the file it should be clear from the 'Description' column what each variable is about and in which units it is in. Ideally this is also mostly celar form the name of the variable. 

```{r models-head, exercise=TRUE}
names(input_table)
```

### Perform a Monte Carlo simulation

Using the model function above, we can perform a Monte Carlo simulation with the `mcSimulation()` function from `decisionSupport`. This function generates distributions of all variables in the input table as well as the specified model outputs (see `return()` function above) by calculating random draws in our defined `example_decision_function()`. Make sure that all the variables in the input table are included in the model (erroneous variables listed there can cause issues with some of the post-hoc analyses). 

The `numberOfModelRuns` argument is an integer indicating the number of model runs for the Monte Carlo simulation. Unless the model function is very complex, 10,000 runs is a reasonable choice (for complex models, 10,000 model runs can take a while, so especially when the model is still under development, it often makes sense to use a lower number).

```{r models-seminar-mcSimulation, exercise=TRUE}
mcSimulation_results <- decisionSupport::mcSimulation(
  estimate = decisionSupport::estimate_read_csv("data/example_input_table.csv"),
  model_function = example_decision_function,
  numberOfModelRuns = 200,
  functionSyntax = "plainNames"
)
```

### Model assessment

For the assessment of the results we will use the `decisionSupport` and many `tidyverse` libraries [@tidyverse2019] including `ggplot2` [@R-ggplot2], `plyr` [@R-plyr] and `dplyr` [@R-dplyr] among others in the [R programming language](https://www.r-project.org/) [@R-base]. 
 
### Plot Net Present Value (NPV) distributions 

We can use the `plot_distributions()` function to produce one of the several plotting options for distribution outputs. This shows us an overlay of the full results of the Monte Carlo model of the decision options, i.e. the expected NPV if we choose to do the intervention `Interv_NPV` or not do the intervention `NO_Interv_NPV`.

```{r models-seminar-plot_distributions, exercise=TRUE}
decisionSupport::plot_distributions(mcSimulation_object = mcSimulation_results, 
                                    vars = c("Interv_NPV", "NO_Interv_NPV"),
                                    method = 'smooth_simple_overlay', 
                                    base_size = 7)

```

We can use the same function to show the distributions of the 'do' `Interv_NPV` and 'do not do' `NO_Interv_NPV` decision scenarios as boxplots. This can be useful when comparing multiple outputs by illustrating the spread of the data resulting from the decision model. Boxplots show the median (central line), the 25^th^ and 75^th^ percentiles (sides of boxes) and any outliers (light circles outside of boxes). 

```{r models-seminar-plot_distributions_boxplot, exercise=TRUE}
decisionSupport::plot_distributions(mcSimulation_object = mcSimulation_results, 
                                    vars = c("Interv_NPV",
                                    "NO_Interv_NPV"),
                                    method = 'boxplot')
```

We can use the same function for the value of the decision (difference in NPV between do and do not do). This can be quite helpful for us since it shows us the outcome distribution of the decision itself. 

```{r models-seminar-plot_distributions_box_dens, exercise=TRUE}
decisionSupport::plot_distributions(mcSimulation_object = mcSimulation_results, 
                                    vars = "NPV_decision_do",
                                    method = 'boxplot_density')
```


#### Cashflow analysis

Here we plot the distribution of annual cashflow over the entire simulated period for the intervention (`n_years`). For this we use the `plot_cashflow()` function which uses the specified cashflow outputs from the `mcSimulation()` function (in our case `Cashflow_decision_do`) to show cashflow over time.

```{r models-seminar-plot_cashflow, exercise=TRUE}
plot_cashflow(mcSimulation_object = mcSimulation_results, cashflow_var_name = "Cashflow_decision_do")
```

#### Projection to Latent Structures (PLS) analysis

We apply a post-hoc analysis to the `mcSimulation()` outputs with `plsr.mcSimulation()` to determine the Variable Importance in the Projection (VIP) score and coefficients of a Projection to Latent Structures (PLS) regression model. This function uses the outputs of the `mcSimulation()` selecting all the input variables from the decision analysis function in the parameter `object` and then runs a PLS regression with an outcome variable defined in the parameter `resultName`. We use the code `names(mcSimulation_results$y)[3]` to select the outcome variable `r names(mcSimulation_results$y)[3]`, which is the third element of the list `y` in our `mcSimulation_results` outputs (this must be a character element). 

```{r models-seminar-pls, exercise=TRUE}
pls_result <- plsr.mcSimulation(object = mcSimulation_results,
                  resultName = names(mcSimulation_results$y)[3], ncomp = 1)
```

We run the `plot_pls()` on the results from `plsr.mcSimulation()` with a number of standard settings. The length of the bars is equal to VIP with a vertical line at '1' on the x-axis indicating a standard cut-off for VIP used for variable selection. The overall plot only shows those variables with a VIP > 0.8, which is the common threshold for variable selection. The colors of the bars represent the positive or negative coefficient of the given input variable with the output variable.

Here we import the input table again to replace the labels for the variables on the y-axis. The input table can include a `label` and `variable` column. The standard labels (from the `variable` column) are usually computer readable and not very nice for a plot. The `plot_pls()` function uses the text in the `label` column as replacement for the default text in the `variable` column.  

```{r models-seminar-plot_pls, exercise=TRUE}
plot_pls(pls_result, input_table = input_table, threshold = 0)
```

#### Value of Information (VoI) analysis

We calculate Value of Information (VoI) analysis with the Expected Value of Perfect Information (EVPI). As we learned in [Lecture 8 on forecasts](#forecasts), EVPI measures the expected opportunity loss that is incurred when the decision-maker does not have perfect information about a particular variable. EVPI is determined by examining the influence of that variable on the output value of a decision model. 

We use the function `data.frame()` to transform the x and y outputs of the `mcSimulation()` function for EVPI calculation. We use the `multi_EVPI()` function to calculate the EVPI for multiple independent variables. For the first_out_var argument we choose `intervention_mngmt_audit_cost` from the input table since this is the first variable after the `NPV` and `cashflow` model outputs, which we would like to exclude from the EVPI analysis.

```{r models-seminar-evpi, exercise=TRUE}
#here we subset the outputs from the mcSimulation function (y) by selecting the correct variables
# choose this carefully and be sure to run the multi_EVPI only on the variables that the you want
mcSimulation_table <- data.frame(mcSimulation_results$x, mcSimulation_results$y[1:3])

evpi <- multi_EVPI(mc = mcSimulation_table, first_out_var = "Interv_NPV")
```

We use the function `plot_evpi()` on the results from `multi_EVPI()` to plot the Expected Value of Perfect Information (EVPI). Here we show the results with the standard settings. The length of the bars is equal to EVPI.

```{r models-seminar-evpi_plot, exercise=TRUE}
plot_evpi(evpi, decision_vars = "NPV_decision_do")
```

Finally, we can use the `compound_figure()` function to provide a single figure for a quick assessment. The can be used to run the full decision assessment for a simple binary decision ('do' or 'do not do').

```{r models-seminar-compound_figure, exercise=TRUE}
compound_figure(mcSimulation_object = mcSimulation_results, 
                input_table = input_table, plsrResults = pls_result, 
                EVPIresults = evpi, decision_var_name = "NPV_decision_do", 
                cashflow_var_name = "Cashflow_decision_do", 
                base_size = 7)

```

### Tasks

In our seminar meeting each group will present the decision and proposed impact pathway model that they will be programming for the rest of the course. See the model section of the talk by Hoa Do in the [Lecture 8 on forecasts](#forecasts) (from around 4:40-6 minutes in) for an example of how to present a model overview in succinct and clear way. The model is presented in the first three minutes of [the talk](https://youtu.be/MlIetFhu87U). 

<!-- make a list of groups in Slack -->